---
author: "Benjamin Meyer, Kenai Watershed Forum"
date: "`r Sys.Date()`"
---

# (APPENDIX) Appendix {.unnumbered}

# Appendix: Data Uplift to EPA WQX

## Introduction

Prior to analysis and interpretation of water quality data, we will ensure that all data that meets QA/QC standards outlined in the project Quality Assurance Project Plan (QAPP) [@kenaiwatershedforum2020] is accessible in the appropriate repository. Water quality data for this project is ultimately destined for the EPA Water Quality Exchange (EPA WQX), formerly EPA STORET.

Section B10 of [@kenaiwatershedforum2020] describes data management details and responsible parties for each step of the data pipeline.in

### 2021 Water Quality Data

Water quality data generated from the Kenai River Baseline Water Quality Monitoring (KRBWQM) program was submitted to the Soldotna office of the Alaska Department of Environmental Conservation in January 2022.

```{r, notes}

# course of action

# uplift 2021 data

# audit existing data in WQX

# attempt to download missing data from  AQWMS, then upload to EPA WQX

# if not available, reshape data from shared drive to uplift to EPA WQX (big task)

```

<br>

## 2021 Baseline Water Quality Data

```{r, other notes, include = F}

# ultimately, would like to set up auto query of EPA WQX so this report updates automatically (github actions or other methods...?); eg see front matter guts of https://geocompr.robinlovelace.net/.

```

